A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2018-12-02 02_Materials_and_Methods.Rmd
one more half sentence about the Mantel test

The continous timeseries of yearwise burial rite proportions for each region allow to calculate timeseries of a measure of cultural distance for each region relation (Figure **TODO**). Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**TODO**). In this context the Squared Euclidian Distance (SED) seems to be a sensible statistic, because it is fairly simple, works with only few assumptions and emphasizes differences through squaring.

@@

To test for correlation between the distances based on the variables *burial type* and *burial construction*, the timeframe of interest was devided into sections of 200 years. For each section of each the two variables, a matrix of the mean SED over time was calculated. The two matrizes of each section can be tested for correlation with the Mantel test, which **TODO** [@mantelDetectionDiseaseClustering1967]. Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test. The mean correlation for each of the $\frac{8*8}{2}-\frac{8}{2}=28$ region-region relations can be plotted in the *burial type*SED-*burial construction*SED space (Figure **TODO**).

>>>>

The continous timeseries of yearwise burial rite proportions for each region allow to calculate timeseries of a measure of cultural distance for each region relation (Figure **TODO**). Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**citation**). In this context the Squared Euclidian Distance (SED) seems to be a sensible statistic, because it is fairly simple, works with only few assumptions and emphasizes differences through squaring.

@@

To test for correlation between the distances based on the variables *burial type* and *burial construction*, the timeframe of interest was devided into sections of 200 years. For each section of each the two variables, a matrix of the mean SED over time was calculated. The two matrizes of each section can be tested for correlation with the Mantel test, which uses permutations of allegedly correlating matrizes to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967]. Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test. The mean correlation for each of the $\frac{8*8}{2}-\frac{8}{2}=28$ region-region relations can be plotted in the *burial type*SED-*burial construction*SED space (Figure **TODO**).
